---
title: "Boca Leek IHW Benchmark"
output: html_document
---

## Preperation
First some cleaning and basic packages.
```{r setup, message = F}
library(magrittr)
library(dplyr)
library(ggpubr)
library(ggplot2)
library(cowplot)
knitr::opts_chunk$set(echo = TRUE)

library(doParallel)
n.cores <- parallel::detectCores()
doParallel::registerDoParallel(cores = min(5, n.cores - 1))
library(doRNG)
set.seed(123)
theme_set(theme_cowplot())
```

Now we need to download my current version of IHW. Make sure the official IHW package is not loaded. In doubt, detach IHW before. Make sure you have the most current version of `remotes`.
```{r load_functions, eval = T}
devtools::load_all()
# library("IHW")
```

## 0.preperation 

### Functions used throughout

This will be the function to evaluate the IHW results. We want to control the False Discovery Rate (FDR) at a prespecified level, typically 0.1. We want to maximize power, e.g. number of discoveries.

### Setting parameters 

Setting some parameters used throughout. 
```{r set_parameters}
m <- 1e4
r <- 200
#r <- 3 # number of monte carlo replicates, increases run time immensely!
alpha <- .1
```

The following parameters are auto-set in the BioConductor implementation. My code is not capable of this yet. 
```{r set_parameters2}
lambda <- Inf

ntrees <- 3L # number of trees per forest
ntaus <- 10
taus <- NULL

nsplit <- 5
maxdepth <- 10 # 3 worked well!
min.node.size <- NULL
nbins <- 3L

folds_fdp_eval <- sample(1:5, m, replace = TRUE)
#dimensions <- seq(from = 2, to =6, by = 2)
dimensions <- seq(from = 1, to =11, by = 1)
```

For the random forest approach, a censoring parameter `tau` is introduced. I will revisit auto setting later. `ntrees` controls the number of trees in each random forest.

## 1. grouped Simulation with signal

```{r}
prop_alt <- function(cov_row) {
  r <- sum(cov_row^2)
  #exp(-5*r)
  #1 / (1 + exp(-cov1))
  #1 / (1 + exp(-1 * (3 * cov_row[1]+cov_row[2] - 5)))
  #browser()
  #ifelse(sum(cov_row^2) <= 0.5, 0.9, 0)
  #ifelse(r <= 0.1, 0.9, 0)
  #1 / (1 + exp(-cov_row[1]))
  #ifelse(sum(cov_row^2)  <= 1, 0.02, 0.4)
  #ifelse(cov_row[1]^2+cov_row[2]^2  <= 1, 0.02, 0.4)
  ifelse(cov_row[1] <= 0.1, 0.9, 0)
  #ifelse(r <= 0.1, 0.9, 0) #This works well
}
```


```{r, eval = T}
sim_combs <- expand.grid(
  m = m,
  dimensions = dimensions,
  seed = seq_len(r)
)

#simple_sim <- foreach(i = seq_len(r)) %dorng% {
simple_sim <- lapply(seq_len(nrow(sim_combs)), function(i) {
  dimension_i <- sim_combs$dimensions[i]
  covariate_i <- matrix(runif(m * dimension_i, 0, 1), nrow = m)
  prop_alt_i <- apply(covariate_i, 1, prop_alt)
  Hs_i <- rbinom(m, size = 1, prop_alt_i)
  
  #mus <- pmax(1.3, sqrt(covariate_i[,c(1,2),drop = F]) %*% c(1,1)*1.5)
  #mu_alphas <- 1/mus
  #pvalue_i <- stats::runif(m)*(1-Hs_i) + stats::rbeta(m, mu_alphas, 1)*Hs_i
  
  pvalue_i <- ifelse(Hs_i,
        rbeta(m, 0.25, 1),
        runif(m)
      )
   
  return(list(covariate = covariate_i, prop_alt = prop_alt_i, Hs = Hs_i, pvalue= pvalue_i))
})
```

```{r, eval =F}
# Generate a 100-dim covariate x
set.seed(0)
m <- 100
n <- 1000
x <- matrix(runif(n * m), n, m)

# Generate the parameters from a conditional two-group
# logistic-Gamma GLM  where pi(x) and mu(x) are both
# linear in x. pi(x) has an intercept so that the average
# of pi(x) is 0.3
inv_logit <- function(x) {exp(x) / (1 + exp(x))}
pi1 <- 0.3
beta.pi <- c(3, 3, rep(0, m-2))
beta0.pi <- uniroot(function(b){
    mean(inv_logit(x %*% beta.pi + b)) - pi1
}, c(-100, 100))$root
pi <- inv_logit(x %*% beta.pi + beta0.pi)
beta.mu <- c(2, 2, rep(0, m-2))
beta0.mu <- 0
mu <- pmax(1, x %*% beta.mu + beta0.mu)

# Generate p-values
H0 <- as.logical(ifelse(runif(n) < pi, 1, 0))
y <- ifelse(H0, rexp(n, 1/mu), rexp(n, 1))
pvals <- exp(-y)

# Run adapt_xgboost
library(adaptMT)
res <- adapt_xgboost(x, pvals, s0 = rep(0.15, n), nfits = 5)
```

```{r, message=FALSE, warning=FALSE, eval = F}
#devtools::install_github("ryurko/adaptMT")
#eval_adap_boost <- lapply(seq_len(nrow(sim_combs)), function(i){
i <- 1
  simple_sim_i <- simple_sim[[i]]
  dimension_i <- sim_combs$dimensions[i]
  seed <- sim_combs$seed[i]
  covariate_i <- simple_sim_i$covariate

  adaptMT::adapt_xgboost_cv	
  adaptMT::adapt_xgboost_cv(
    covariate_i,
    simple_sim_i$pvalue,
  verbose = list(print = FALSE,
                 fit = FALSE,
                 ms = FALSE),
  #piargs = args_search,
  #muargs = args_search,
  n_folds = 5#,
  #niter_ms = 10,
  #nms = as.integer(2),
  #s0 = rep(0.05, nrow(bmi_esnps_data))
  )

  
  rjs_i <- IHWStatsPaper::adapt_mtp(simple_sim_i$pvalue, covariate_i, alpha = 0.1, formula_rhs = "~.")
  Hs_i <- simple_sim_i$Hs
  evaluated_i <- data.frame(pow = sum(rjs_i * Hs_i) / max(1, sum(Hs_i)), 
                            FDP = sum(rjs_i * (1 - Hs_i)) / max(1, sum(rjs_i)),
                            dimension = dimension_i)
  evaluated_i
#})

eval_adap <- do.call(rbind, eval_adap)
eval_adap$strat <- "AdaPT"
```

```{r, message=FALSE, warning=FALSE, eval = F}
eval_adap <- lapply(seq_len(nrow(sim_combs)), function(i){
  simple_sim_i <- simple_sim[[i]]
  dimension_i <- sim_combs$dimensions[i]
  seed <- sim_combs$seed[i]
  covariate_i <- simple_sim_i$covariate

  rjs_i <- IHWStatsPaper::adapt_mtp(simple_sim_i$pvalue, covariate_i, alpha = 0.1, formula_rhs = "~.")
  Hs_i <- simple_sim_i$Hs
  evaluated_i <- data.frame(pow = sum(rjs_i * Hs_i) / max(1, sum(Hs_i)), 
                            FDP = sum(rjs_i * (1 - Hs_i)) / max(1, sum(rjs_i)),
                            dimension = dimension_i)
  evaluated_i
})

eval_adap <- do.call(rbind, eval_adap)
eval_adap$strat <- "AdaPT"
```

```{r, message=FALSE, warning=FALSE, eval = T}
eval_quantile_simple <- lapply(seq_len(nrow(sim_combs)), function(i){
  #simple_sim, function(simple_sim_i) {
  simple_sim_i <- simple_sim[[i]]
  dimension_i <- sim_combs$dimensions[i]
  seed <- sim_combs$seed[i]
  
  covariate_i <- simple_sim_i$covariate
  ihw_quantile <- ihw(simple_sim_i$pvalue, covariate_i, alpha = 0.1, strat = "cut", nbins = nbins, lambda = Inf, folds = folds_fdp_eval, null_proportion = T) 

  rjs_i <- IHW::rejected_hypotheses(ihw_quantile)
  Hs_i <- simple_sim_i$Hs
  evaluated_i <- data.frame(pow = sum(rjs_i * Hs_i) / max(1, sum(Hs_i)), 
                            FDP = sum(rjs_i * (1 - Hs_i)) / max(1, sum(rjs_i)),
                            dimension = dimension_i)
  evaluated_i
})

eval_quantile_simple <- do.call(rbind, eval_quantile_simple)
eval_quantile_simple$strat <- "IHW-Quantile"
```


```{r, message=FALSE, warning=FALSE, eval = T}
eval_bocaleek_simple <- lapply(seq_len(nrow(sim_combs)), function(i){
  #simple_sim, function(simple_sim_i) {
  simple_sim_i <- simple_sim[[i]]
  dimension_i <- sim_combs$dimensions[i]
  seed <- sim_combs$seed[i]
  covariate_i <- simple_sim_i$covariate

  ihw_bocaleek <- ihw(simple_sim_i$pvalue, covariate_i,
    alpha = 0.1, strat = "BocaLeek", ntrees = ntrees,
    taus = taus, ntaus = ntaus, maxdepth = maxdepth, min.node.size = min.node.size, lambda = Inf, nsplit = nsplit, null_proportion = T)

  rjs_i <- IHW::rejected_hypotheses(ihw_bocaleek)
  Hs_i <- simple_sim_i$Hs
  evaluated_i <- data.frame(pow = sum(rjs_i * Hs_i) / max(1, sum(Hs_i)), 
                            FDP = sum(rjs_i * (1 - Hs_i)) / max(1, sum(rjs_i)),
                            dimension = dimension_i)
  evaluated_i
})

eval_bocaleek_simple <- do.call(rbind, eval_bocaleek_simple)
eval_bocaleek_simple$strat <- "IHW-Forest"
```

```{r, message=FALSE, warning=FALSE, eval = T}
eval_bh_simple <- lapply(simple_sim, function(simple_sim_i) {
  rjs_i <- p.adjust(simple_sim_i$pvalue, n = m, method = "BH") <= 0.1
  
  Hs_i <- simple_sim_i$Hs
  evaluated_bh_i <- data.frame(pow = sum(rjs_i * Hs_i) / max(1, sum(Hs_i)), 
                               FDP = sum(rjs_i * (1 - Hs_i)) / max(1, sum(rjs_i)))
  evaluated_bh_i
})
eval_bh_simple <- do.call(rbind, eval_bh_simple)
eval_bh_simple$strat <- "BH"
eval_bh_simple <- merge(eval_bh_simple,
                        data.frame(dimension = dimensions))
```

```{r, message=FALSE, warning=FALSE}
eval_all <- rbind(eval_quantile_simple, eval_bocaleek_simple, eval_bh_simple)# eval_bocaleek_noise  eval_bh_simple, eval_quantile_simple, eval_bocaleek_simple eval_adap, 

eval_all$strat <- factor(eval_all$strat,
    levels = c('BH',  'IHW-Quantile',"IHW-Forest"),ordered = TRUE)#'quantile', "BocaLeek more Noise" 'AdaPT',
    
eval_all <- eval_all %>%
  group_by(dimension, strat) %>%
  summarize(
    FDR = mean(FDP, na.rm = TRUE),
    Power = mean(pow, na.rm = TRUE),
    n_monte_carlo = sum(!is.na(FDP)),
    pow_se = sd(pow, na.rm = TRUE) / sqrt(n_monte_carlo)
  ) %>%
  ungroup()

 eval_all_FDR <- ggplot(eval_all, aes_string(x = "dimension", y = "FDR", shape = "strat", col = "strat")) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept = alpha, linetype = 2) +
  guides(color=guide_legend(title="Method", legend.text=element_text(size=4)),
         shape =  "none")+
   scale_x_continuous(breaks= scales::pretty_breaks())+
   scale_color_manual(values = c("#00AFBB", "#52854C", "#FC4E07"))# fill=guide_legend(nrow=2,byrow=TRUE)
 
eval_all_power <- ggplot(eval_all, aes_string(x = "dimension", y = "Power", shape = "strat", col = "strat")) +
  geom_line() +
  geom_point()  +
  ylab("Power")+
  guides(color=guide_legend(title="Method", legend.text=element_text(size=4)))+
  scale_x_continuous(breaks= scales::pretty_breaks())+
   scale_color_manual(values = c("#00AFBB", "#52854C", "#FC4E07"))

g_combined<- ggarrange(eval_all_FDR, eval_all_power, nrow = 1, widths = c(1, 1), common.legend = TRUE, legend = "bottom") 

#g_combined <- annotate_figure(g_combined, top = text_grob("ifelse(sum(cov_row[1]^2) <= 1, 0.9, 0)", 
#               color = "red", face = "bold", size = 14))
```

```{r}
ggplot2::ggsave("~/R/IHW-1/figures/benchmark.png", g_combined,
                width = 5, height = 3.25)
g_combined
```

