---
title: "maf_simulation"
format: html
editor: visual
bibliography: references.bib
---

```{r echo=FALSE}
library("dplyr")
library("GGally")
library(here)
options(bitmapType ="cairo")
```

## Estimate simulation hyper parameters based on real data

```{r, eval = FALSE}
chr <- 21
chr_full_data_grubert <- loadRData(paste0("/g/scb/zaugg/zaugg/hQTL/GenotypeData/snps", chr, ".filtered.allGT.txt.rda"))

```

Data driven parameter for $s_i$

```{r, eval = FALSE}
s_prob <- mean(chr_full_data_grubert[[1]])/2
s_prob
```

Same number of patients

```{r, eval = FALSE}
nsample <- ncol(chr_full_data_grubert)
nsample
```

Read fitted modell

```{r, eval = FALSE}
chr_pvalues_grubert <- data.table::fread(here(paste0("data/hqtl_chrom1_chrom2/results/cisQTLs_H3K27AC_chr",chr,".txt")))
head(chr_pvalues_grubert)
```

```{r, eval = FALSE}
betha <- mean(chr_pvalues_grubert$beta)
```

sanity check

```{r, eval = FALSE}
chr_pvalues_grubert$`p-value`[1]
2*(1-pt(abs(chr_pvalues_grubert$`t-stat`[1]), df = 74))
```

## Simulation

We are following the notation in [@shabalin2012]. For a fixed SNP and a fixed histone modification peak, let $s_i\in \{0,1,2\}$ be the frequency of the minor allele and $g_i \in \mathbb{R}_{\geq 0}$ be the level of histone modification measured in individual $i \in \{1,\ldots, n\}$. Let $H \sim \mathop{Bernouilli}(1-\pi_0)$ indicate, if there actually is an association between $s_i$ and $g_i$. We model the relationship as follows:

$$
\begin{aligned} &g_i =\alpha+(H\beta)  s_{i}+\epsilon_{i}.\end{aligned}
$$

with $s \sim \operatorname{Bin}(2,p)$ and $\epsilon \sim \mathcal{N}\left(0, \sigma^2\right)$. We will be testing

$$
H_0: (H\beta) =0 \quad \text { against } \quad H_1: (H\beta) \neq 0.  
$$

Hyperparameters for simulation

```{r}
ntrials = 1000
nsample = 76
pi0 = 0.9
```

```{r}
#alpha should not play any role
simulate_data <- function(H, nsample = 76, s_prob = 0.26, alpha = 50, betha = 30, sigma = 2){
  data_sim <- data.frame(
    #save parameters (unobserved)
    s_prob = s_prob,
    alpha = alpha,
    betha = betha, #true
    sigma = sigma,
    nsample_i = seq_len(nsample),
    
    #simulate s and noise
    s = rbinom(n = nsample, size = 2, s_prob),
    noise = rnorm(n = nsample, sd = sigma)
  )

  data_sim <- data_sim %>% 
    mutate(histone_modification = alpha + betha * s*H  + noise)
  
  #only histone_modification and s observable
  return(data_sim)
}
```

```{r}
H <- rbinom(ntrials, size = 1, 1-pi0)
data_sim <- lapply(H, simulate_data)

head(data_sim[[1]])
```

## Minor allele frequency

Let $\bar{s} := \frac{1}{n}\sum_{i=1}^n s_{i}$. Then the minor allele frequency can be defined as

$$
f:=\frac{1}{2 n} \min \left(\sum_{i=1}^n s_i, 2 n-\sum_{i=1}^n s_i\right)=\frac{1}{2} \min (\bar{s}, 2-\bar{s}) \in[0,0.5]
$$

    That is, of the $2n$ alleles present in $n$ individuals, what the proportion of the minor allele at the pre-specified SNP is.

```{r}
minor_allele_frequencies_sim <- sapply(data_sim, function(data_sim_i){
  0.5* min(mean(data_sim_i$s), 2- mean(data_sim_i$s))
})
```

## t-test

We test

$$
H_0: (H\beta) =0 \quad \text { against } \quad H_1: (H\beta) \neq 0.  
$$

Our test statistic is

$$
t_{\text {score }}=\frac{\widehat{\beta}}{S E_{\widehat{\beta}}},
$$

where $\widehat{\beta}$ is the least square estimator for $H\beta$ and $S E_{\widehat{\beta}}$ is the associated standard error. We are following the notation from B7.13 in [@czado2011]. Under the null and assuming $\epsilon_{1},\ldots , \epsilon_{n} \sim \text { i.i.d. } \mathcal{N}\left(0, \sigma^2\right)$ we have $t_{\text {score }} \sim \mathcal{T}_{n-2}$.

```{r}
sim_res <- lapply(data_sim, function(data_sim_i){
  t_test_lm = lm(histone_modification ~ 1 + s, data=data_sim_i)
  s_line_t_test_lm = summary(t_test_lm)$coefficients[2,]
  
  s_line_t_test_lm = data.frame(as.list(s_line_t_test_lm))
  
  #calculate minor_allele_frequency
  s_line_t_test_lm$minor_allele_frequency = 0.5* min(mean(data_sim_i$s), 2- mean(data_sim_i$s))
  
  s_line_t_test_lm$s_variance <- var(data_sim_i$s)
  
  s_line_t_test_lm
})

sim_res <- bind_rows(sim_res)
#sim_res <- sim_res %>% rename(pvalue = `Pr...t..`)
colnames(sim_res)[4] <- "pvalue"
sim_res$H <- as.factor(H)
head(sim_res)
```

## Rationale for MAF

We don't think that the minor allele frequency $f$ informs about the hypothesis $H$. We think that under the alternative $H=1$ it is informative for our multiple testing procedure in the following way:

Higher MAF $f =\frac{1}{2}\min \left( \bar{s},2- \bar{s}\right)$

$\Rightarrow$ Higher predictor variation $\sqrt{\sum_{i=1}^n\left(s_i-\bar{s}\right)^2}$ (intuitive, both are measures of variation of $s$)

$\Rightarrow$ Smaller variation of $\widehat{\beta}$ estimate $$
S E_{\widehat{\beta}}=\frac{\sqrt{\frac{1}{n-2} \sum_{i=1}^n\left(g_i-\bar{g}\right)^2}}{\sqrt{\sum_{i=1}^n\left(s_i-\bar{s}\right)^2}}
$$

$\Rightarrow$ Higher test statistic $$
t_{\text {score }}=\frac{\widehat{\beta} }{S E_{\widehat{\beta}}}
$$

$\Rightarrow$ Smaller p-value

$$P = 2(1-F_{H_0}(t_{\text {score }}))$$

::: callout-note
Question: 1) Is above interpretation correct? 2) What happens under the null above?
:::

## Diagnostic plots

```{r}
ggpairs(sim_res, 
        mapping=ggplot2::aes(colour = H, alpha = 0.3),
        columns= c(5,6,2,3,4), 
        title="correlogram with ggpairs()",
        lower=list(combo=wrap("facethist", binwidth=1))
      ) 

```

## Pvalue histograms

```{r}
groups_by_filter <- function(covariate, nbins, ties.method="random", seed=NULL){
  if (!is.null(seed) && ties.method=="random"){
    #http://stackoverflow.com/questions/14324096/setting-seed-locally-not-globally-in-r?rq=1
    tmp <- runif(1)
    old <- .Random.seed
    on.exit( { .Random.seed <<- old } )
    set.seed(as.integer(seed)) 
  }
	rfs <- rank(covariate, ties.method=ties.method)/length(covariate)
	as.factor(ceiling( rfs* nbins))
}
```

```{r}
ggplot(sim_res, 
       aes(x = pvalue,
           fill = H)) +
  geom_histogram(boundary = 0) 
```

```{r}
ggplot(sim_res, 
       aes(x = pvalue,
           fill = H)) +
  geom_histogram(boundary = 0) +
  facet_wrap(groups_by_filter(sim_res$minor_allele_frequency, 8), nrow = 2)
```
