---
title: "Boca Leek IHW Benchmark"
output: html_document
---

## Preperation
First some cleaning and basic packages.
```{r setup, message = F}
library(magrittr)
library(dplyr)
library(ggpubr)
library(ggplot2)
library(cowplot)
knitr::opts_chunk$set(echo = TRUE)

library(doParallel)
n.cores <- parallel::detectCores()
doParallel::registerDoParallel(cores = min(5, n.cores - 1))
library(doRNG)
set.seed(123)
theme_set(theme_cowplot())
```

Now we need to download my current version of IHW. Make sure the official IHW package is not loaded. In doubt, detach IHW before. Make sure you have the most current version of `remotes`.
```{r load_functions, eval = T}
# remotes::install_github("FridljDa/IHW", subdir = "bioconductor")
# summaryDir <- "/g/huber/users/fridljand/R/HIGH/data/14_summary"
# censDir <- "/Volumes/fridljand/R/HIGH/data/05_demog"
# devtools::install_github("Huber-group-EMBL/covariate-powered-cross-weighted-multiple-testing", subdir = "IHWStatsPaper")
# devtools::load_all("/Users/default/Google Drive/currentDocumants/Studium/Master/3.Semester/Masterarbeit/IHW")#
# devtools::load_all("/Users/default/Google Drive/currentDocumants/Studium/Master/3.Semester/Masterarbeit/RFCDE")
devtools::load_all() #

library("IHW")

devtools::load_all("/Users/default/Google Drive/currentDocumants/Studium/Master/3.Semester/Masterarbeit/IHW/huber_github/IHWStatsPaper")
# library("RFCDE")
```

## 0.preperation 

### Functions used throughout

This will be the function to evaluate the IHW results. We want to control the False Discovery Rate (FDR) at a prespecified level, typically 0.1. We want to maximize power, e.g. number of discoveries.
```{r}
calc_ihw <- function(strat, sim = NULL, Xs = NULL, Ps = NULL, alpha, folds, ntrees, tau, lambda) {
  if (!is.null(sim)) {
    Xs <- lapply(sim, function(sim_i) sim_i$Xs)
    Ps <- lapply(sim, function(sim_i) sim_i$Ps)
  }
  ihw_res <- foreach(Xs_i = Xs, Ps_i = Ps) %dopar% {
    ihw_res_i <- ihw(Ps_i, Xs_i, alpha = alpha, folds = folds, strat = strat, ntrees = ntrees, tau = tau, lambda = lambda)
  }
  return(ihw_res)
}
```

### Setting parameters 

Setting some parameters used throughout. 
```{r set_parameters}
m <- 10000
r <- 2 # number of monte carlo replicates, increases run time immensely!
alpha <- .1
```

The following parameters are auto-set in the BioConductor implementation. My code is not capable of this yet. 
```{r set_parameters2}
nfolds <- 3
folds <- sample(1:nfolds, m, replace = TRUE)
lambda <- Inf
```

For the random forest approach, a censoring parameter `tau` is introduced. I will revisit auto setting later. `ntrees` controls the number of trees in each random forest.
```{r set_parameters3}
tau <- 0.7 # censoring parameter for BocaLeek estimator
ntrees <- 100 # number of trees per forest
```

## 4. But does it work better in univariate case?
Lets return to a simple, univariate example. The official referencemanual (http://bioconductor.org/packages/release/bioc/manuals/IHW/man/IHW.pdf) works with following data.
```{r generatedata_univ, message=FALSE}
Xs <- lapply(1:r, function(i) {
  runif(m, min = 0, max = 2.5) # covariate
})
Hs <- lapply(1:r, function(i) {
  rbinom(m, 1, 0.1) # hypothesis true or false
})
Ps <- lapply(1:r, function(i) {
  Xs_i <- Xs[[i]]
  Hs_i <- Hs[[i]]
  Z_i <- rnorm(m, Hs_i * Xs_i) # Z-score
  Ps_i <- 1 - pnorm(Z_i) # pvalue
})
```

Lets run IHW forest and the quantile slicing on this: 
```{r run_IHW_univ, message=FALSE, eval = T}
ihw_univ_quantile <- foreach(Xs_i = Xs, Ps_i = Ps) %dopar% {
  ihw(Ps_i, Xs_i, alpha = alpha, folds = folds, strat = "quantile", lambda = lambda)
}
```

## 2. BetaMix Example

Now we create the betamix sample data according to Section 5.2 in https://doi.org/10.1111/rssb.12411 with two-dimensional covariates
```{r, eval = T}
betamix_sim_combs <- expand.grid(
  mu_slope = seq(1),
  seed = seq_len(r)
)

betamix_sim <- foreach(i = 1:nrow(betamix_sim_combs)) %dorng% {
  betamix_sim_i <- IHWStatsPaper::beta_unif_sim(m = m, mus_slope = betamix_sim_combs$mu_slope[i], one_sided_tests = FALSE, prob_one_sided = 0.25)
  betamix_sim_i$Xs <- as.matrix(betamix_sim_i$Xs)
  colnames(betamix_sim_i$Xs) <- c("X1", "X2")
  betamix_sim_i
}
```
```{r}
ihw_betamix_bocaleek <- foreach(betamix_sim_i = betamix_sim) %dopar% {
  Xs_i = betamix_sim_i$Xs
  Ps_i = betamix_sim_i$Ps
  ihw_forest_bocaleek_i <- ihw(Ps_i, Xs_i, alpha = alpha, folds = folds, strat = "BocaLeek", ntrees = ntrees, tau = tau, lambda = lambda)
}
length(ihw_betamix_bocaleek[[1]]@df[["pvalue"]])
```

## 3. BetaMix with noise     
Lets add some uninformative noise. If this still works, this would be extremely convenient for the end user. Furthermore, so many dimensions definitely necessitates the use of regression trees. Due to the exponential increase of the number of bins with the dimension, quantile slicing would be computationally infeasible. The ability to incorporate high-dimensional covariates is the most convincing advantage of the random forest approach. 
```{r}
Xs_noise <- foreach(betamix_sim_i = betamix_sim) %dopar% {
  Xs_i = betamix_sim_i$Xs
  noise <- matrix(runif(m * 5), nrow = m)
  Xs_noise_i <- cbind(Xs_i, noise)
  colnames(Xs_noise_i) <- paste0("X",1:7)
  Xs_noise_i <- as.matrix(Xs_noise_i)
  Xs_noise_i
}

```


```{r, message=FALSE, warning=FALSE, eval = T}
ihw_betamix_noise_bocaleek <- foreach(Xs_noise_i = Xs_noise, betamix_sim_i = betamix_sim) %dopar% {
  Ps_i <- betamix_sim_i$Ps
  ihw(Ps_i, Xs_noise_i, alpha = alpha, folds = folds, strat = "BocaLeek", ntrees = ntrees, tau = tau, lambda = lambda)
}

length(ihw_betamix_noise_bocaleek[[1]]@df[["pvalue"]])
```

#does it work
```{r create data}
mus_slope <- 1.5
prob_one_sided <- 0.25

Xs <- lapply(1:r, function(i) {
  Xs_i_i <- matrix(runif(m * 2, 0, 1), ncol = 2)
  colnames(Xs_i_i) <- c("X1", "X2")
  Xs_i_i
})

Hs <- lapply(Xs, function(Xs_i) {
  pi1s <- ifelse(Xs_i[, 1]^2 + Xs_i[, 2]^2 <= 1, 0.02, 0.4)
  Hs_i <- stats::rbinom(m, size = 1, prob = pi1s)
  Hs_i
})

Ps <- mapply(function(Xs_i, Hs_i) {
  mus <- pmax(1.3, sqrt(Xs_i) %*% c(1, 1) * mus_slope)
  mu_alphas <- 1 / mus
  Ps_i <- stats::runif(m) * (1 - Hs_i) + stats::rbeta(m, mu_alphas, 1) * Hs_i
  Ps_i
}, Xs, Hs)
```

```{r, message=FALSE, warning=FALSE}
ihw_betamix_noise_slic <- foreach(Xs_i = Xs, Ps_i = Ps) %dopar% {
  noise <- matrix(runif(m * 5), nrow = m)
  Xs_i <- cbind(Xs_i, noise)
  ihw(Ps_i, Xs_i, alpha = alpha, folds = folds, strat = "BocaLeek", ntrees = ntrees, tau = tau, lambda = lambda)
}

```
