---
title: "Boca Leek IHW Benchmark"
author: "Daniel Fridljand"
date: "`r Sys.Date()`"
#output: rmarkdown::html_vignette
header-includes:
   - \usepackage{bbm}
output: html_document
---

## Preperation
First some cleaning and basic packages.
```{r setup, message = F}
library(magrittr)
library(dplyr)
library(ggpubr)
library(ggplot2)
library(cowplot)
library(RColorBrewer)
knitr::opts_chunk$set(echo = TRUE)

set.seed(123)
theme_set(theme_cowplot())
options(bitmapType ="cairo")
```

## Colors and method names used throughout 

```{r}
options(bitmapType ="cairo")
# (from https://github.com/dill/beyonce)
beyonce_colors <- c("#b72da0", "#7c5bd2", "#0097ed","#00c6c3",
                    "#9cd78a", "#f7f7a7", "#ebab5f", "#e24344",
                    "#04738d")#,"#d8cdc9")
beyonce_colors[6] <- c("#dbcb09") # thicker yellow
pretty_colors <- beyonce_colors[c(2,1,3:5)]
alpha = 0.1
```

```{r}
method_names <- c("BH", "Clfdr-oracle", "Clfdr-EM", "AdaPT", "IHW-Storey-BetaMix", "IHW-BH-Grenander")
method_colors <- c("Black", beyonce_colors[c(3,4, 6, 2, 1)])
names(method_colors) <- method_names
method_shapes <- c(20, 0,1, 2, 12, 13)
names(method_shapes) <- method_names

```

# Noise simulation

\begin{eqnarray}
& X_i \sim U[0,1]^d \\
&\pi_1(X) = \mathbf{1}(X^1 \leq 0.5)0.9\\
&H_i \mid X_i \sim \text{Bernoulli}(\pi_1(X_i))\\
&P_i \mid H_i = 0, X_i \sim U[0,1]\\
&P_i \mid H_i = 1, X_i \sim Beta(0.25, 1)
\end{eqnarray}
Note parameter $d$ in the first line. Hereafter, dimension = $d$.


## Load result files
```{r}
noise_sim_res <-readRDS(file="../precomputed_results/noise_sim.Rds")
noise_sim_res <- noise_sim_res %>% mutate(method = forcats::fct_recode(method,
"IHW-Quantile"= "IHW-quantile",
"IHW-Forest" ="IHW-forest" ))
```

```{r, message=FALSE, warning=FALSE}
noise_sim_res <- noise_sim_res %>%
  group_by(dimension, method) %>%
  summarize(
    FDR = mean(FDP, na.rm = TRUE),
    Power = mean(pow, na.rm = TRUE),
    #Power_lower = quantile() #TODO
    n_monte_carlo = sum(!is.na(FDP)),
    pow_se = sd(pow, na.rm = TRUE) / sqrt(n_monte_carlo)
  ) %>%
  ungroup()
```

## Plot FDR
```{r}
noise_sim_res_FDR <- ggplot(noise_sim_res, aes_string(x = "dimension", y = "FDR", shape = "method", col = "method")) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept = alpha, linetype = 2) +
  guides(color=guide_legend(title="Method", legend.text=element_text(size=4)),
         shape =  "none")+
   scale_x_continuous(breaks= scales::pretty_breaks())
  #+ xlab(expression(bar(beta)))
  #+ #scale_y_log10(breaks = c(0.003, 0.01, 0.025, 0.05, 0.1, 0.2, 0.4))
   #scale_color_manual(values = c("#00AFBB", "#52854C", "#FC4E07"))# fill=guide_legend(nrow=2,byrow=TRUE)
noise_sim_res_FDR
```
## Plot Power

```{r}
noise_sim_res_power <- ggplot(noise_sim_res, aes_string(x = "dimension", y = "Power", shape = "method", col = "method")) +
  geom_line() +
  geom_point()  +
  ylab("Power")+
  guides(color=guide_legend(title="Method", legend.text=element_text(size=4)))+
  scale_x_continuous(breaks= scales::pretty_breaks())#+
   #scale_color_manual(values = c("#00AFBB", "#52854C", "#FC4E07"))
noise_sim_res_power
```

## Combine and save

```{r}
g_combined<- ggarrange(noise_sim_res_FDR, noise_sim_res_power, nrow = 1, widths = c(1, 1), 
                       common.legend = TRUE, legend = "bottom") 

#g_combined <- annotate_figure(g_combined, top = text_grob("ifelse(sum(cov_row[1]^2) <= 1, 0.9, 0)", 
#               color = "red", face = "bold", size = 14))
g_combined
ggplot2::ggsave("../figures/noise_sim.png", g_combined,
                width = 5, height = 3.25)

```
# Small region simulation

\begin{eqnarray}
& X_i \sim U[0,l] \\
&\pi_1(X) = \mathbf{1}(X^1 \leq 1)0.9\\
&H_i \mid X_i \sim \text{Bernoulli}(\pi_1(X_i))\\
&P_i \mid H_i = 0, X_i \sim U[0,1]\\
&P_i \mid H_i = 1, X_i \sim Beta(0.25, 1)
\end{eqnarray}

Note parameter $l$ in the first line. Hereafter, length = $l$.

## Load result files
```{r}
small_region_sim_res <-readRDS(file="../precomputed_results/small_region_sim.Rds")
small_region_sim_res <- small_region_sim_res %>% mutate(method = forcats::fct_recode(method,
"IHW-Quantile"= "IHW-quantile",
"IHW-Forest" ="IHW-forest" ))
```

```{r, message=FALSE, warning=FALSE}
small_region_sim_res <- small_region_sim_res %>%
  group_by(length, method) %>%
  summarize(
    FDR = mean(FDP, na.rm = TRUE),
    Power = mean(pow, na.rm = TRUE),
    #Power_lower = quantile() #TODO
    n_monte_carlo = sum(!is.na(FDP)),
    pow_se = sd(pow, na.rm = TRUE) / sqrt(n_monte_carlo)
  ) %>%
  ungroup()
```

## Plot FDR
```{r}
small_region_sim_res_FDR <- ggplot(small_region_sim_res, aes_string(x = "length", y = "FDR", shape = "method", col = "method")) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept = alpha, linetype = 2) +
  guides(color=guide_legend(title="Method", legend.text=element_text(size=4)),
         shape =  "none")+
   scale_x_continuous(breaks= scales::pretty_breaks())+
  scale_y_log10()
  #+ xlab(expression(bar(beta)))
  #+ #scale_y_log10(breaks = c(0.003, 0.01, 0.025, 0.05, 0.1, 0.2, 0.4))
   #scale_color_manual(values = c("#00AFBB", "#52854C", "#FC4E07"))# fill=guide_legend(nrow=2,byrow=TRUE)
small_region_sim_res_FDR
```

## Plot Power

```{r}
small_region_sim_res_power <- ggplot(small_region_sim_res, aes_string(x = "length", y = "Power", shape = "method", col = "method")) +
  geom_line() +
  geom_point()  +
  ylab("Power")+
  guides(color=guide_legend(title="Method", legend.text=element_text(size=4)))+
  scale_x_continuous(breaks= scales::pretty_breaks())+
 scale_y_log10()
#breaks = c(0.003, 0.01, 0.025, 0.05, 0.1, 0.2, 0.4)
   #scale_color_manual(values = c("#00AFBB", "#52854C", "#FC4E07"))
small_region_sim_res_power
```

## Combine and save

```{r}
g_combined<- ggarrange(small_region_sim_res_FDR, small_region_sim_res_power, nrow = 1, widths = c(1, 1), 
                       common.legend = TRUE, legend = "bottom") 

g_combined
```

```{r}
ggplot2::ggsave("../figures/small_region_sim.png", g_combined,
                width = 5, height = 3.25)
```

