---
title: "IHW simple Benchmark"
output: html_document
---


Loading packages.
```{r setup, message = F}
library(magrittr)
library(dplyr)
knitr::opts_chunk$set(echo = TRUE)
set.seed(1)
```

```{r, eval = T}
JuliaCall::julia_setup(JULIA_HOME = "/Applications/Julia-1.6.app/Contents/Resources/julia/bin/")
JuliaCall::julia_command("cd(\"/Users/default/Google Drive/currentDocumants/Studium/Master/3.Semester/Masterarbeit/Code/IndependentHypothesisWeightingTrees.jl\")")
JuliaCall::julia_source("/Users/default/Google Drive/currentDocumants/Studium/Master/3.Semester/Masterarbeit/Code/IndependentHypothesisWeightingTrees.jl/example/wrapper.jl")
```

Generate simulation from https://support.bioconductor.org/p/90005/#9137847. 

```{r}
m <- 1e5
nfolds <- 3
folds <- sample(1:nfolds, m, replace = TRUE)

cov1 <- runif(m, -1, 1)
cov2 <- runif(m, -1, 1)
prop_alt <- 1 / (1 + exp(-1 * (3 * cov1 + 2 * cov2 - 5)))
pvalue <- ifelse(rbinom(m, size = 1, prop_alt),
  rbeta(m, 0.25, 1),
  runif(m)
)
noise_dim <- 2
covariate <- cbind(cov1, cov2)
noise <- matrix(rnorm(noise_dim*m), m)
covariate_noise <- cbind(covariate,noise)
covariate_noise <- cbind(cov1, noise)
```

Loading my working version of IHW using random forest. 
```{r load_functions, eval = T, message=F, results='hide'}
devtools::load_all()
#devtools::install_github("FridljDa/RFCDE")
#library("IHW")
#devtools::load_all("../../../RFCDE")
library("RFCDE")
```


Set parameters for IHW Forest
```{r set_parameters3}
tau <- 0.5 # censoring parameter for BocaLeek estimator
#nbasis <- 10L
#ntrees <- 10L # number of trees per forest
#nbins <- 10L
lambda <- Inf

ntrees <- 10L # number of trees per forest
nbins <- 10L
nbasis <- 10L
```

The RFCDE variant takes longer, but is more powerful:

```{r,  message=F}
ihw_cut <- ihw(pvalue, covariate_noise, folds =folds, alpha = 0.1, strat = "cut",  lambda = lambda, nbins = nbins)
rejections(ihw_cut)
```
```{r}
ihw_Julia <- ihw(runif(1000), as.matrix(runif(1000)), alpha = 0.1, strat = "Julia", ntrees = 10, lambda = Inf, nbins = 10, nbasis = 10)
```

```{r}
ihw_Julia <- ihw(pvalue, covariate_noise, folds =folds,alpha = 0.1, strat = "Julia", ntrees = ntrees, lambda = lambda, nbins = nbins, nbasis = nbasis)
rejections(ihw_Julia)
```

```{r,  message=F, eval = T}
ihw_Julia <- ihw(pvalue, covariate_noise, folds =folds,alpha = 0.1, strat = "Julia", ntrees = ntrees, lambda = lambda, nbins = nbins, nbasis = nbasis)
rejections(ihw_Julia)
```

Now we run the BocaLeek variant. Note, that we use the actual cov1 and cov2, not the binned version.
```{r,  message=F}
ihw_bocaleek <- ihw(pvalue, covariate_noise,folds =folds, alpha = 0.1, strat = "BocaLeek", ntrees = ntrees, tau = tau, nbins = nbins, lambda = lambda)
rejections(ihw_bocaleek)
```


The RFCDE variant takes longer, but is more powerful:
```{r,  message=F, eval = T}
ihw_rfcde_haar <- ihw(pvalue, covariate_noise, folds =folds,alpha = 0.1, strat = "RFCDE", ntrees = ntrees, nbins = nbins, lambda = lambda, nbasis = nbasis, basis_system = "Haar") 
rejections(ihw_rfcde_haar)
```
```{r,  message=F, eval = T}
ihw_rfcde_cosine <- ihw(pvalue, covariate_noise, folds =folds,alpha = 0.1, strat = "RFCDE", ntrees = ntrees, nbins = nbins, lambda = lambda, nbasis = nbasis, basis_system = "cosine") 
rejections(ihw_rfcde_cosine)
```

TODO
```{r,  message=F, eval = T}
ihw_rfcde_legendre <- ihw(pvalue, covariate_noise, folds =folds,alpha = 0.1, strat = "RFCDE", ntrees = ntrees, nbins = nbins, lambda = lambda, nbasis = nbasis, basis_system = "Legendre") 
rejections(ihw_rfcde_legendre)
```

So the early results from the forest variant already look encouraging. However, we still have lots of ideas on how to further improve the method.
Regarding FDR: FDR control is pretty much guaranteed from the theoretical point of view by the mathematical framework Nikos provided. We have further verified this on a few simulations. 

```{r, load_airway, message = F, warning=FALSE}
library("DESeq2")
data("airway", package = "airway")
dds <- DESeq2::DESeqDataSet(se = airway, design = ~ cell + dex) %>% DESeq()
deRes <- as.data.frame(results(dds))
```

setting some parameters
```{r, message = F}
m <- nrow(deRes)
folds <- sample(1:nfolds, m, replace = TRUE)
```

```{r,  message=F, eval = F}
ihw_Julia <- ihw(deRes$pvalue, deRes$baseMean, folds =folds,alpha = 0.1, strat = "Julia", ntrees = ntrees, lambda = lambda, nbins = nbins, nbasis = nbasis)
rejections(ihw_Julia)
```

```{r}
ihw_airway_quantile <- ihw(deRes$pvalue, deRes$baseMean, alpha = 0.1, folds = folds, lambda = lambda, strat = "quantile")
rejections(ihw_airway_quantile)
```
