---
title: "IHW simple Benchmark"
output: html_document
---


Loading packages.
```{r setup, message = F}
library(magrittr)
library(dplyr)
library("IHW")
knitr::opts_chunk$set(echo = TRUE)

set.seed(1)
```

Generate simulation from https://support.bioconductor.org/p/90005/#9137847. 

```{r}
n <- 1e5
cov1 <- runif(n, -1, 1)
cov2 <- runif(n, -1, 1)
prop_alt <- 1 / (1 + exp(-1 * (3 * cov1 + 2 * cov2 - 5)))
pvalue <- ifelse(rbinom(n, size = 1, prop_alt),
  rbeta(n, 0.25, 1),
  runif(n)
)

dat <- data.frame(
  cov1 = cut(cov1, breaks = quantile(cov1, 0:4 / 4), include.lowest = TRUE),
  cov2 = cut(cov2, breaks = quantile(cov2, 0:4 / 4), include.lowest = TRUE),
  prop_alt,
  pvalue
)

dat <- dat %>% mutate(group = factor(paste(as.integer(cov1), as.integer(cov2), sep = "-")))
```

First, we try normal  normal BH:
```{r}
padjBH <- p.adjust(dat$pvalue, method = "BH")
sum(padjBH <= 0.1)
```

Next, we run the BioConductor version of IHW with the binned covariates. 
```{r, eval = F}
# run IHW on the grid
res <- IHW::ihw(pvalue ~ group,
  data = dat,
  alpha = 0.1,
  covariate_type = "nominal"
)
rejections(res)
```

Loading my working version of IHW using random forest. 
```{r load_functions, eval = T, message=F, results='hide'}
devtools::load_all("../../../IHW-1")
#devtools::install_github("FridljDa/RFCDE")
#devtools::load_all("IHW")
devtools::load_all("../../../RFCDE")
library("IHW")
```


Set parameters for IHW Forest
```{r set_parameters3}
tau <- 0.7 # censoring parameter for BocaLeek estimator
n_basis <- 5
ntrees <- 10 # number of trees per forest
lambda <- Inf

pvalue <- dat$pvalue
covariate <- cbind(cov1, cov2)
groups <- dat$group
```

```{r,  message=F}
ihw_Julia <- ihw(pvalue, covariate, alpha = 0.1, strat = "Julia", ntrees = ntrees, lambda = lambda, n_basis = 10)
rejections(ihw_Julia)
```

Now we run the BocaLeek variant. Note, that we use the actual cov1 and cov2, not the binned version.
```{r,  message=F}
ihw_bocaleek <- ihw(pvalue, covariate, alpha = 0.1, strat = "BocaLeek", ntrees = ntrees, tau = tau, lambda = lambda)
rejections(ihw_bocaleek)
```

The RFCDE variant takes longer, but is more powerful:
```{r,  message=F}
ihw_rfcde <- ihw(pvalue, covariate, alpha = 0.1, strat = "RFCDE", ntrees = ntrees, lambda = lambda, n_basis = 10)
rejections(ihw_rfcde)
```


So the early results from the forest variant already look encouraging. However, we still have lots of ideas on how to further improve the method.
Regarding FDR: FDR control is pretty much guaranteed from the theoretical point of view by the mathematical framework Nikos provided. We have further verified this on a few simulations. 
