---
title: "RFCDE"
author: "Taylor Pospisil"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{RFCDE}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Random forests is a common non-parametric regression technique which
performs well for mixed-type data and irrelevant covariates, while
being robust to monotonic variable transformations. RFCDE fits random
forest models optimized for nonparametric conditional density
estimation.

## Example
```{r load_functions, echo = F, message = F, eval = F}
devtools::load_all()
```



```{r set parameters}
m <- 20000
nfolds <- 5
nbins <- 5
node_size <- floor(m / (3 * nbins)) # todo smaller trees are ok for forest
```

```{r generate data}
Xs <- runif(m, min = 0, max = 2.5) # covariate
Hs <- rbinom(m, 1, 0.6) # hypothesis true or false
Zs <- rnorm(m, Hs * Xs) # Z-score
folds <- sample(1:nfolds, m, replace = TRUE)
Ps <- 1 - pnorm(Zs) # pvalue
Xs <- as.matrix(Xs)
# todo rename Ps to be more consistent
```

### Training

Trees are recursively partitioned to minimize the CDE loss

$$ \int \int (\hat{f}(x \mid z) - f(x \mid z))^{2} dz dP(x) $$

This is efficiently calculated using an orthogonal series
representation of the conditional densities. The resolution of this
representation is controlled by `n_basis`.


### Prediction

We use the forest structure to determine weights for a weighted kernel
density estimate. The `predict` function evaluates this density on the
provided grid.

#group_by_forest2 <- function(pvalues, covariates, folds, nfolds, nbins, ntrees = 10, seed=NULL)

```{r}
n_trees <- 10
mtry <- 4
n_basis <- 15
groups <- matrix(0, nrow = nrow(Xs), ncol = n_trees)

i <- 1
for (i in seq_len(nfolds)) {
  # Ps_holdout_fold <- Ps[folds == i]
  # Xs_holdout_fold <- Xs[folds == i, ]
  holdout_fold <- which(folds != i)
  Ps_other_folds <- Ps[folds != i]
  Xs_other_folds <- Xs[folds != i, ]

  forest_other_folds <- RFCDE(
    x_train = Ps_other_folds, z_train = Xs_other_folds, n_trees = n_trees, mtry = mtry,
    node_size = node_size, n_basis = n_basis
  )

  forest_other_folds$rcpp$set_leaves_id()

  for (j in holdout_fold) {
    groups[j, ] <- forest_other_folds$rcpp$traverse_forest(Xs[j, ]) #TODO as matrix
  }
}

groups <- as.data.frame(groups)
groups[] <- lapply(groups, as.factor)
colnames(groups) <- paste0("group", seq_along(groups))
```
